---
title: "Consumer Product Time-Series Forecasting"
author: "Ilse Paniagua"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fpp3)
library(tsibble)
library(readr)
library(prettydoc)
```

This RMarkdown document shows tidy tools for time series data using the `tsibble`, `fable`, and `feasts`packages for time series data frames, tidy forecasting, and feature extraction and statistics.

I used the following data sources.

* Cookies: https://fred.stlouisfed.org/series/APU0000702421
U.S. Bureau of Labor Statistics, Cookies, Chocolate Chip, Per Lb. (453.6 Gm) in U.S. City Average [APU0000702421], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/APU0000702421, January 14, 2020. 

* Flour: 
https://fred.stlouisfed.org/series/APU0000701111
U.S. Bureau of Labor Statistics, Flour, White, All Purpose, Per Lb. (453.6 Gm) in U.S. City Average [APU0000701111], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/APU0000701111, January 14, 2020

* Bread: 
https://fred.stlouisfed.org/series/APU0000702111
U.S. Bureau of Labor Statistics, Bread, White, Pan, Per Lb. (453.6 Gm) in U.S. City Average [APU0000702111], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/APU0000702111, January 14, 2020. 

References:

* Hyndman, R.J., & Athanasopoulos, G. (2019) Forecasting: principles and practice, 3rd edition, OTexts: Melbourne, Australia. OTexts.com/fpp3).

# Data Preparation

```{r message=FALSE}
#Reading each dataset as a tsibble object

flour <- as_tsibble(read_csv("flour_1980-2019.csv"), index=DATE)
bread <- as_tsibble(read_csv("bread_1980-2019.csv"), index=DATE)
cookies <- as_tsibble(read_csv("cookies_1980-2019.csv"), index=DATE)
```

I combine all three datasets and do some variable renaming. Because forecasting requires that all missing values be explicit rather than implict, I fill in possible time gaps with "NA".

```{r}
#Joining dataframes and renaming variables
ts <- cbind(flour, bread, cookies)

#Renaming and dropping variables
ts <- select(ts, date = DATE, price_flour = APU0000701111, -DATE1, price_bread = APU0000702111, -DATE2, price_cookies = APU0000702421)

#Changing date into year-month format
#ts$date <- as_date(ts$date)
ts$date <- yearmonth(ts$date)

#Filling in missing values (none applicable)
fill_gaps(ts, .full=FALSE)
```

I creating individual files for each time series. Each file has 480 observations (12 months * 40 years)

```{r}
flour <- select(ts, date, price_flour)
bread <- select(ts, date, price_bread)
cookies <- select(ts, date, price_cookies)
```

# Plotting the data

Plotting the time series of each dataset.

```{r message=FALSE}
#Plot with all time series
ggplot(data=ts, aes(x=date)) +
  geom_line(aes(y=price_bread), color="blue") +
  geom_line(aes(y=price_flour), color="red") +
  geom_line(aes(y=price_cookies), color="black") +
  labs(y="Price",
       x="Date",
       title="U.S. City Average price of cookies, bread, flour 1980-2019 (per lb)",
       caption="Source: U.S. Bureau of Labor Statistics"
       ) +
  theme_light()

#Cookies plot
ggplot(data=ts, aes(x=date)) +
  geom_line(aes(y=price_cookies), color="black") +
  labs(y="Price",
       x="Date",
       title="U.S. City Average price of cookies 1980-2019 (per lb)",
       caption="Source: U.S. Bureau of Labor Statistics"
       ) +
  theme_light()


#Bread plot
ggplot(data=ts, aes(x=date)) +
  geom_line(aes(y=price_bread), color="blue") +
  labs(y="Price",
       x="Date",
       title="U.S. City Average price of bread 1980-2019 (per lb)",
       caption="Source: U.S. Bureau of Labor Statistics"
       ) +
  theme_light()

#Flour plot
ggplot(data=ts, aes(x=date)) +
  geom_line(aes(y=price_flour), color="red") +
  labs(y="Price",
       x="Date",
       title="U.S. City Average price of flour 1980-2019 (per lb)",
       caption="Source: U.S. Bureau of Labor Statistics"
       ) +
  theme_light()

```

* Cookies:

The price of cookies increased from about 1.5 dollars per pound in 1980 to over 3.5 dollars per pound in 2019. The ts displays a positive trend and seasonality.


* Flour:

The price of lour increased from about 0.2 cents in January 1980 to about 0.4 cents in December 2019.It has some seasonal fluctuations and exhibits an overal positive trend. There was a sharp increase in the price of flour from 2008 to 2009.

* Bread:

The bread ts is trended and seasonal. Prices increased from 0.5 cents 1980 to over $1.5 2012, and decreased slightly from 2012 to 2019 to about 1.33 dollars.


## Decomposition

All three time series exhibit a strong seasonability and a positive trend. The remainder term increases in variability as the time series progresses.

```{r}
cookies %>% STL(price_cookies ~ season(window = Inf)) %>% autoplot()
```


```{r}
flour %>% STL(price_flour ~ season(window = Inf)) %>% autoplot()
```

```{r}
bread %>% STL(price_bread ~ season(window = Inf)) %>% autoplot()
```

# Defining a model

## Cookies

We obtain forecasts for the next two years using both ARIMA and ETS models. The `ETC()` function selects the model parameters by minimising the AIC_c. The `ARIMA` function in R uses a variation of the Hyndman-Khandakar algorithim to obtain an ARIMA model.

```{r}
cookies %>%
  model(
    ets = ETS(price_cookies),
    arima = ARIMA(price_cookies)
  ) %>%
  forecast(h = "2 years") %>%
  autoplot(cookies) +
    labs(title = "Two-year forcasts for price of cookies",
       y="Price",
       x="Date") +
  theme_light()
```

The forecasts using the two models look very similar. But which is more accurate?

To compare the forecast accuracy of ETS and ARIMA models, we use data before December 2014 as our training data and January 2015-December 2019 as our test data.

We cannot rely on minimizing the AIC_c to compare between ETC and ARIMA models because they are in different model classes and the likelihood is computed in different ways. Instead, we use time-series cross validation to assess model accuracy.

```{r}
#Five-year forecasts 2015-2019
fit.cookies <- cookies %>%
  filter(date < yearmonth("2014 Dec")) %>%
  model(
    ETS(price_cookies),
    ARIMA(price_cookies)) %>%
  forecast(h= "5 years")
 
#Obtaining accuracy statistics
acc.cookies <- fit.cookies %>% accuracy(cookies)

acc.cookies
```

ARIMA provides more accurate forecasts than ETS for the cookies time series according to all metrics.

The selected model for the cookies ts is an ARIMA(0,1,3)(0,0,1) with drift.

```{r}
cookies %>%
  model(ARIMA(price_cookies)) %>% report()
```

## Flour

For the flour ts, there is a wider discrepancy between ETS and ARIMA forecasts. ETS estimates higher prices for flour than ARIMA.

```{r}
flour %>%
  model(
    ets = ETS(price_flour),
    arima = ARIMA(price_flour)
  ) %>%
  forecast(h = "2 years") %>%
  autoplot(flour) +
  labs(title = "Two-year forcasts for price of flour",
       y="Price",
       x="Date") +
  theme_light()
```

We again assess model accuracy using time-series cross validation.

```{r}
fit.flour <- flour %>%
  filter(date < yearmonth("2014 Dec")) %>%
  model(
    ETS(price_flour),
    ARIMA(price_flour)) %>%
  forecast(h= "5 years")
 
acc.flour <- fit.flour %>% accuracy(flour)

acc.flour
```

ARIMA again outperforms ETCs for the flour data. An ARIMA(0,1,0)(2,0,0) model will be used for forecasting.

```{r}
flour %>%
  model(ARIMA(price_flour)) %>% report()
```

## Bread

For bread, the ETS and ARIMA models produce similar forecasts.

```{r}
bread %>%
  model(
    ets = ETS(price_bread),
    arima = ARIMA(price_bread)
  ) %>%
  forecast(h = "2 years") %>%
  autoplot(bread) +
    labs(title = "Two-year forcasts for price of bread",
       y="Price",
       x="Date") +
  theme_light()
```

We assess model accuracy using time-series cross validation.

```{r}
fit.bread <- bread %>%
  filter(date < yearmonth("2014 Dec")) %>%
  model(
    ETS(price_bread),
    ARIMA(price_bread)) %>%
  forecast(h= "5 years")
 
acc.bread <- fit.bread %>% accuracy(bread)

acc.bread
```
ETS outperforms ARIMA for this dataset. We use an ETS(M,A,N) for this dataset.

```{r}
bread %>%
  model(ETS(price_bread)) %>% report()
```


# Residual diagnostics

A good forecasting method will yield residuals that are:

1. Uncorrelated
2. Have a zero mean

If the forecasts are correlated, there is additional information that should be used in computing forecasts. If the residuals do not have a mean of zero, they will be biased.

For each time-series, we produce a time plot, ACF plot, and histogram of the residuals for our models. We will also perform Ljung-Box tests for autocorrelation.

For the Ljung-Box tests, we use lag parameters as outlined by Rob J Hyndman (https://robjhyndman.com/hyndsight/ljung-box-test/).

He suggests:
* For non-seasonal time series, use h=min(10,T/5)
* For seasonal time series, use h=min(2m,T/5),

where m is the length of the seasonality and T is the length of the time-series.

Because all our series our seasonal, we will use a lag of 8 (40/5).

## Cookies

```{r message=FALSE, warning=FALSE}
cookies%>%
  model(ETS(price_cookies)) %>%
  gg_tsresiduals()
```

The graphs show that the model seems to account for all possible information. The mean of the residuals is close to zero and the variation of the residuals stays close to constant across the historical data. The histogram suggests that the residuals are normal. With the exception of two lags that show slight significance, he lack of correlation in the ACF plot suggets that forecasts are acceptable.

We perform a Ljung-Box test to assess whether autocorrelations are significantly different from white noise. The null hypothesis is that the residual series is white noise; therefore failure to reject (large p-value) means that our model is appropriate.

The results of the Ljung-Box test are not significant. Therefore, we conclude that the residuals are not distinguishable from a white noise series.

```{r}
cookies %>%
  model(ARIMA(price_cookies)) %>%
  augment() %>%
  select(.resid) %>%
  features(.resid, ljung_box, lag=8, dof=5)
```
## Flour

We perform residual diagnostics for the flour dataset using an ARIMA model.

```{r message=FALSE, warning=FALSE}
flour %>%
  model(ARIMA(price_flour)) %>%
  gg_tsresiduals()
```


```{r}
flour %>%
  model(ARIMA(price_flour)) %>%
  augment() %>%
  select(.resid) %>%
  features(.resid, ljung_box, lag=8, dof=5)
```
The residuals for flour are more problematic. There is evidence that the residuals do not have a mean of zero and have some periods of variability, particularly from 2005-2010 (roughly when the series underwent a level change). The residuals also fail the Ljung-Box test for the ARIMA model. This shows that the model has not perfectly captured the information in the data. However, this is the best model that we have evaluated.

## Bread

We repeat this exercise for the bread time-series. We had selected an ETS model.

```{r message=FALSE, warning=FALSE}
bread %>%
  model(ETS(price_bread)) %>%
  gg_tsresiduals()
```

```{r}
bread %>%
  model(ETS(price_bread)) %>%
  augment() %>%
  select(.resid) %>%
  features(.resid, ljung_box, lag=8, dof=6)
```

The residuals for the bread dataset seem to have some large outliers based on the time plot and the histogram. A few observations seem to be autocorrelated based on the ACF plot. The residuals bread dataset fail the Ljung-Box test based on an alpha=0.5 value.

Now I will develop a Shiny app to showcase these results!